### 2020-12-17 Thesis-Chapter6-ThreatModel.tex
Editorial improvements

This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
proposal is specific to \acl{DAR} for mobile devices. Due to the length of its title, this proposal will be referred to
by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
application and use, though it is not the focus of this analysis, and nothing is assumed about it in particular. Note
that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
\ac{EA} proposal, but use the term in the spirit of their paper.

@@

``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The first step to
answering those questions is to establish goals for the former and a list of threats in anticipation of the latter. The
\ldawmsr proposal describes a threat model against which it was designed. This model takes cues from that one but is
intended to be more generally applicable for any \ac{DAR} \ac{EA} proposal.

@@

The goals of an \ac{EA} cryptosystem are taken from the metrics of \mychap{chap-arguments}. Their justification stems
from the preceding arguments. Each goal is listed according to priority, though no goal has absolute priority over the
others.

@@

\modelitem{Security}{The system should minimize security risk with reference to the current state of the art. Risk is
considered with respect to both individual devices and the entire class of devices.}

\modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
authoritarian regimes. This includes technologically enforcing the fundamental right to privacy but upon warranted
search pursuant to the rule of law.}

\modelitem{Transparency}{The system should be auditable by the public to allow for verification that its power is being
used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}

@@

\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem and the
distribution of costs associated with its development and administration. Even though this is a security-oriented threat
model, architectural decisions determine the system's economics, and economics determines the system's ultimate
feasibility.}

@@

\ac{EA} threat models include the full cast of traditional threat actors.

@@

\modelitem{Criminals: ordinary, organized, cyber}{Criminals threaten the \ac{EA} system by using it to access plaintext
without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical
access but only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but
considerably more advanced technical capabilities, exerting full control over all interfaces. Cyber criminals have a
wide spectrum of motivations and capabilities. Their significant contribution to the threat landscape is that they will
target the backend systems that operate the \ac{EA} system in order to steal and likely sell whatever access
capabilities they can find.}

\modelitem{Insider threat}{The insider threat involves someone within the \ac{EA} recovery process, perhaps at the law
enforcement agency or a vendor, that seeks to abuse the system for their own purposes. They may have the same
motivations and skills of any other criminal category, but they have the advantage of privileged access at some stage in
the process.}

@@

\modelitem{Authoritarian regimes}{States that do not respect lawful process will try to compel administrators of an
\ac{EA} system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking
capability.}

@@

To make sense of an \ac{EA} proposal in the face of the stated goals and such a formidable list of threat actors, one
must include two non-traditional threats. To do so is to step back and view the situation with a broader perspective
than security experts usually take.

@@

\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
The end user is not the most important thing. This is a crucial point.

All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of this value system is some most
important ``thing.'' Traditional device security threat models define this ``thing'' as the end user. This leads to good
security priorities and honors the user's rights and autonomy.

To stray from this rule is dangerous, but is already frequent practice. Software that forces updates values the vendor's
judgement over the user's (even if it is a good idea). Software that enforces Digital Rights Management values
corporations' data rights over the user's (sometimes to extremes \cite{eff_2020}). Software that does not encrypt user
communications at all values the service provider's ability to scrape personal information over the user's privacy.
Software that enables \ac{EA} values lawful investigatory process over the user's ultimate privacy. Intentional
subversion of the user's will is not a comfortable thought, but it is a frequent reality, and the reality of \ac{EA}.

In the broader threat model, the platform abuser uses encryption to hide illicit activities. He is a threat to public
safety.}

\modelitem{Hawkish lawmakers}{If one's very highest priorities are security and privacy for society as a whole, then
aggressive lawmakers that have the power to \ii{mandate} arbitrary data access represent perhaps the largest threat of
all. Crypto-anarchists can encrypt their data above the compromised layer all they want, but above-the-table
organizations and most individuals who use default products will abide by U.S. law.

@@

\modelitem{Encyption workarounds}{Encryption workarounds are always in play for essentially any threat actor at any
time. Discussion for defending against those is out of scope.}

@@

\modelitem{Supply chain attacks}{It is necessary to protecting device hardware and software supply chains in order to
prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
software are considered to be as-intended, though they are assumed to contain the standard amount of bugs and
side-channels.}

\modelitem{Manipulation of internal hardware}{Hyper-advanced hardware analysis and manipulation is a real threat from
foreign intelligence agencies. Defending against these attacks is a separate problem.}

\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
warrant is valid. Additionally, all forms of surveillance, even when protected from \ac{masssurv}, require oversight.
Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}

@@

Answering ``what are we building'' and comparing it with the state of the art requires an understanding of the state of
the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
\ldawmsR. As this analysis is limited to a specific proposal, it is also limited to this reference architecture.
\myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to Apple's public
documentation \cite{apple_2020}.

@@

This \ac{DFD}, and all that follow, is at a specific level of abstraction, simplified in the spirit of ``all models are
wrong, but some are useful.'' The \acp{DFD} are intended to be useful for communication and threat elicitation only, but
are accurate in that capacity.

As the diagram illustrates, several steps occur between the user entering her \ac{PIN} and unlocking of the device. The
operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines it with the device-specific
Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is finally handed to
the Encryption Module to perform encryption and decryption at a hardware level between the operating system and actual
encrypted storage.

The \ac{UID} is not depicted in key storage because it is actually burned into the secure processor silicon, and cannot
be read by software. The Volume key is stored encrypted by the Class key. When a user updates his \ac{PIN}, the
encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors never
handle any keys or secret information besides the \ac{PIN}, and due to inline encryption in the Encryption Module,
persistent storage never handles any \ac{plaintext} data.

@@

\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{EA} requests, and a vendor
mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time vaulting,
authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}; see the original paper
for full detail \cite{savage_lawful_2018}.

@@

Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
secure hardware offers, and \ldawmsr demonstrates this capabilities perfectly. The greatest of these capabilities are
physical possession requirements and strong transparency. Physical possession requirements strongly mitigate mass
surveillance and, as Savage puts it, ``provides a more intuitive compatibility with common under-standings of the
government’s reasonable law enforcement powers (e.g., the ability to seize physical property under court order) than
more information-centric approaches, which may appear covert by comparison'' \cite{savage_lawful_2018}. Transparency is
also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
detectable changes, mitigating both \ac{masssurv} and common criminal abuse.

One clear weakness in the design is the shared secret between the device and the manufacturer used in the authorization
protocol. The key may be stored on the device, but one can access it if they have this shared secret. This is very much
like trusted-party key escrow where the vendor is the trusted party, only they store the authentication key instead of
the unlock key itself. Due to this similarity, one of the biggest \ac{EA} problems of centralized risk is not fully
addressed. This is discussed further in \mysec{sec-prop-threats}.

\ldawmsR's full data flow is illustrated in \myfig{fig-dfd-ldawmsr}. The proposal offers a few variations of the
self-escrow and authorization protocols. This figure uses the asymmetric key pair approach for Class key encryption in
the Secure Enclave, which has stricter hardware requirements; however, Savage offers a symmetric key variant that
achieves the same purposes. The asymmetric key pair approach was chosen here because it is semantically clearer and
easier to illustrate.

@@

The lawful access use case operates as follows: law enforcement obtains a warrant to conduct a search, including the
suspect's mobile device. They obtain the phone, disassemble it to reveal the unexposed \ac{EA} hardware interface, and
begin the time-vaulted access request. After the lockout period is completed, the device offers its Device ID only. The
law enforcement agency's digital forensics department digitally signs the Device ID and sends it to the device vendor,
alongside the warrant and accompanying proof of legal authorization. The vendor's access compliance department audits
the legal documents and submits the digitally signed Device ID to a secure computing environment such as a \ac{HSM}. The
secure environment authenticates the digital signature, looks up the device's private Device Seal Key and authorization
token, and encrypts them with the law enforcement agency's public key. These encrypted artifacts are returned to the
agency, which decrypts and submits them to the device through the dedicated interface. First is the authorization token,
which is hashed and compared to the stored hash. If that values match, the device accepts and uses the private Device
Seal Key to decrypt the Class key, which was escrowed after being encrypted but the public Device Seal Key. Once the
Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for forensic investigation.

This concludes an summarized overview of the system as described in the proposal. However, in a real system, the entire
information lifecycle matters. In order to establish such a system, vendors and law enforcement must establish
mechanisms to populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} is a
\ac{DFD} for these scenarios.

@@

The proposal has been illustrated. What threats does it introduce? Recall from \mysec{sec-crypto-basics} that security
can be expressed in the properties of authentication, integrity, non-repudiation, confidentiality, availability, and
authorization. Threats are often expressed as the negative of these properties: spoofing, tampering, repudiation,
information disclosure, \acl{DOS}, and elevation of privileged \cite{shostack_threat_2014}. This section
analyzes \ldawmsr using these categories. Threats that fit under multiple categories, such as tampering for the effect
of repudiation, are listed only once.

@@

Spoofing threats are listed in \mytab{table-spoofing}. Notably, \ii{S2} is Eran Tromer's attack \cite{tromer_2018} on
computer scientist Ray Ozzie's related \ac{DAR} \ac{EA} protocol called CLEAR \cite{ozzie_2018}. Savage acknowledges the
influence that CLEAR had on \ldawmsR, and specifically included a note that while this proposal is technically
vulnerable to the same threat, the mitigations listed here would help, and furthermore the likelihood of such an attack
is low and the difficulty high.

@@

Tampering threats are listed in \mytab{table-tampering}. \ii{S2} relies on tampering similar to \ii{T1}.

@@

Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed here, all
attacks aimed at gaining unauthorized access to a device constitute information disclosure threats. Most spoofing
attacks are aimed at disclosing information intended for another party.

@@

How much risk does \ldawmsr introduce? The answer lies in a range depending on how many mitigations are implemented.
Some threats introduce inherently low risk due to their difficult nature and low pay off. Others are higher risk but can
be reasonably mitigated. Some threats specifically target transparency, and are difficult to address without trust.
Finally, there are some risks that are irreducible. Each of classes are discussed in turn.

\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low risk either
because they fail to produce useful results, they can be reasonably mitigated, or they are a threat that is already
present in state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and
mitigate the risk posed by these threats as part of the design process, but they do not represent fatal flaws to a
design. Except in combination with other threats, they are not worthy more concern here.

\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are each higher risk threats to the
security, trustworthiness, and effectiveness of the proposal, but can be addressed through a set of related mitigations:
require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device to be
physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities the
lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly is a technical measure that entirely
precludes certain attacks; requiring the device to be on LEA premises is a procedural measure that prevents attackers
from manipulating LE into achieving their own ends; limiting the number of LE facilities that can participate prevents
LE impersonation attacks. Together, they make the proposal considerably harder to attack.

\ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
legal process and cooperation of several parties. But what if the legal process involves secret courts, secret
investigations, and secrecy orders on the vendor \cite{shamsi_2011}? Destructive disassembly or the weaker hardware
transparency mechanism provide non-repudiation that the device was unlocked, if and only if one has physical access to
the device. Technically enforced transparency into device access requests and responses would grant non-repudiation at a
broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR} \ac{EA} proposals do
\cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately, building this
feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.

Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, each representing in slightly different ways the
disclosure of unlock artifact data, the DID-DSK database. This is the truly difficult challenge---centralizing so much
sensitive data makes it an extremely high value target for foreign intelligence and cyber criminals. The vendor is in a
position where they need to protect this database from the most advanced attackers in the world while regularly updating
and accessing it. \Acp{HSM} and strong security policies would make that a difficult task, but not impossible, and one
has to assume that the data would leak eventually. What would leaked unlock artifacts mean? It depends who steals it.
Criminals would use it for personal and financial gain; foreign intelligence and authoritarian regimes would use it to
spy on their enemies. In the worst-case scenario, anyone with some specialized tools and a willingness to break the law
could get into any device under the scheme, provided they destructively disassemble it and wait through the time vault.

@@

This chapter began with a list of goals by which \ac{DAR} \ac{EA} proposals can be measured, including security,
protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section I measure
\ldawmsr against each of these goals.

% Consider goals, threat actors, and loop-back to out of scope items, because they aaaaactually do matter.

\paragraph*{Security} Security is the main topic of this chapter, and has already been discussed thoroughly. \ldawmsr
introduces significant risk in the DID-DSK database itself, but does a good job in many other respects, especially if
the recommended mitigations are implemented. It goes from very hard to attack, when everything is working as intended,
to simple to attack, when the database is leaked. However, even when it becomes trivial to attack, the impact is limited
to devices in physical possession and cannot be hidden from the device user.

\paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
authoritarian regimes. This proposal does an excellent job of protecting against \ac{masssurv} due to its physicality.
Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior
to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After
a leak, these regimes could get into any device they want. If one lived in a country that was willing to take and
destroy their devices to read their contents without legal due process, this would be an unwelcome technology.

\paragraph*{Transparency} \ldawmsr excels at transparency on the device, but fails at transparency to the public. Due to
destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed using this method
is irrefutable. However, this is only possible to confirm if the device is available. If it is stolen and unlocked by
criminals or is lawfully accessed under secretive conditions, both the user and the public have no way of knowing. This
leaves open the possibility for government abuse---though again with the caveat that physical possession requirements
mean it cannot be abused without a trace.

\paragraph*{Law enforcement utility} This proposal involves a few tradeoffs between security and law enforcement
utility. The physical possession, time vault, and limited number of authorized law enforcement laboratory requirements
each insert a barrier to usability. The physical possession requirement is far too important for security and civil
liberties to sacrifice for utility. Furthermore, law enforcement, and search and seizure law generally, is geared
towards physical evidence already. The time vault prevents against time-sensitive threats, though destructive
disassembly (not part of the original proposal) may be enough to cover these cases as well, so the time vault could be
decreased without significant impact. Limiting the number of parties able to use the system mitigates spoofing attacks;
local law enforcement would have to cooperate with one of these parties, such as a regional \ac{FBI} office. (This
really just shifts the spoofing threat up a level, though the shift is appropriate.) In the end, law enforcement still
has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end; the onus is on the government to fund
the \ac{FBI} operation and cultivate cooperation between local and federal agencies.

>>>>

This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
proposal is specific to \acl{DAR} for mobile devices. Due to the length of its title, this proposal will be referred to
by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by a legal framework regulating its
application and use, though the legal framework is not the focus of this analysis.

% Note that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
% emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
% \ac{EA} proposal, but use \ac{EA} in ... [that spirit].

@@

``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The process of
answering these questions begins with setting goals, identifying threat actors, and declaring scope. The \ldawmsr
proposal describes a threat model against which the proposal was designed. This threat model takes cues from the
\ldawmsr model, but is intended to be more generally applicable for any \ac{DAR} \ac{EA} proposal.

@@

The goals of an \ac{EA} cryptosystem are taken from the five metrics used to judge \ac{EA} alternatives and variations
in \mychap{chap-arguments}. Each goal is listed according to priority, though no goal has absolute priority over the
others.

@@

\modelitem{Security}{The system should minimize security risk relative to the current baseline. Risk is considered with
respect to both individual devices and the entire class of devices.}

\modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
authoritarian regimes. This includes technologically enforcing the fundamental right to privacy outside the context of
warranted search under the rule of law.}

\modelitem{Transparency}{The system should be auditable by the public to allow for verification that its power is being
used responsibly. In the case of \ac{DAR}, access should be apparent to the data owner.}

@@

\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem and the
distribution of costs associated with its development and administration. Even though this is a security-oriented threat
model, a system's design affects its economics, and its economics determines its ultimate feasibility.}

@@

\ac{EA} threat models include the full slate of traditional threat actors.

@@

\modelitem{Criminals: ordinary, organized, cyber}{Criminals threaten the \ac{EA} system by using it to access plaintext
without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical access but
only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but considerably
more advanced technical capabilities, including full control over all interfaces. Cyber criminals have a wide variety of
of motivations and capabilities. They will target the backend systems that operate the \ac{EA} system in order to abuse
or sell whatever access capabilities they acquire.}

\modelitem{Insider threat}{Insider threats involve attackers with roles inside the \ac{EA} recovery process who seek to
abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
provider. They may have the same motivations and skills of any other criminal category, but they have the advantage of
privileged access at some stage in the process.}

@@

\modelitem{Authoritarian regimes}{Governments that do not adhere to rule of law will try to compel administrators of an
\ac{EA} system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking
capability.}

@@

Two non-traditional threats must be considered to make sense of an \ac{EA} proposal in the face of the stated goals and
such a formidable list of threat actors. Understanding these threats requires viewing security with a broader than usual
perspective.

@@

\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
Therefore, the end user's will is not the platform's top priority.\\

All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is one top
priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
decisions and honors the user's rights and autonomy.\\

It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the vendor's update
policy over the user's preferences. Digital Rights Management software values corporations' data rights over the user's
(sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's ability to collect
personal information over the user's privacy. \Acl{EA} software values lawful investigation over the user's absolute
privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it is a frequent
reality.\\

In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is a threat to
public safety.}

\modelitem{Hawkish lawmakers}{In the pursuit of security and privacy for society as a whole, aggressive lawmakers that
have the power to \ii{mandate} arbitrary data access present perhaps the largest threat of all. Crypto-anarchists will
encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
use default products will abide by U.S. law.\\

@@

\modelitem{Encyption workarounds}{Encryption workarounds, such as finding a copy of plaintext or guessing a password,
are always a risk. Defending against such workarounds is not the role of the \ac{EA} system itself.}

@@

\modelitem{Supply chain attacks}{It is necessary to protecting device hardware and software supply chains in order to
prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
software are assumed to be uncompromised, aside from the standard amount of bugs and side-channels.}

\modelitem{Manipulation of internal hardware}{Foreign intelligence agencies may have the capability to perform
hyper-advanced hardware analysis and manipulation. Defending against these attacks is not the role of the \ac{EA} system
itself.}

\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes legitimacy of the legal process used to
obtain the warrant. All forms of government search require oversight. Transparency via auditability, preferably ensured
via technical mechanism, should be a goal for \ac{EA} proposals. However, rule of law is an absolute prerequisite for
which no \ac{EA} mechanism can compensate.}

@@

Understanding the \ldawmsr proposal and comparing it to current mobile device encryption technology requires an
understanding of current technology. Apple iPhones are among the most secure consumer mobile phones and serve as the
point of reference for \ldawmsR. Because this analysis is limited to a specific proposal, it is also limited to this
reference architecture. In \myfig{fig-dfd-iphone}, a \ac{DFD} depicts the iPhone's unlock and decryption process
according to Apple's public documentation \cite{apple_2020}.

@@

This \ac{DFD}, and all that follow, presents the system at a specific level of abstraction. The system representation is
simplified in the spirit of ``all models are wrong, but some are useful.'' The \acp{DFD} exclude some details, but are
accurate for the purposes of communication and threat elicitation.

As the diagram illustrates, several steps occur between the \ac{PIN} being entered and the device being unlocked. The
operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines the \ac{PIN} with the
device-specific Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is
passed to the Encryption Module. The Encryption Module performs encryption and decryption at a hardware level between
the operating system and the encrypted storage.

The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
software. The Volume key is encrypted by the Class key before it is stored. When a user updates his \ac{PIN}, the
encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
not handle any keys or secret information besides the \ac{PIN}. Due to inline encryption in the Encryption Module,
persistent storage does not handle \ac{plaintext} data.

@@

\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{EA} requests, and a
vendor-mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time
vaulting, authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}.

@@

Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
secure hardware offers---namely, physical possession requirements and strong transparency. The physical possession
requirement mitigates the risk of \ac{masssurv}. As Savage adds, it also ``provides a more intuitive compatibility with
common under-standings of the government’s reasonable law enforcement powers (e.g., the ability to seize physical
property under court order) than more information-centric approaches, which may appear covert by comparison''
\cite{savage_lawful_2018}. Transparency is also rooted in the physicality of the approach, wherein the device cannot be
unlocked without making irreversible and detectable changes, reducing the risk of both \ac{masssurv} and common criminal
abuse.

One clear weakness in the design is its reliance on the secret used in the authorization protocol shared between the
device and the manufacturer. Although the key is stored on the device, anyone could access the key if they know the
shared secret. This is similar to trusted-party key escrow. The vendor is the trusted party, and they store the
authorization key rather than the encryption key itself. Due to this similarity, the problem of centralized risk is not
fully addressed. This is discussed further in \mysec{sec-prop-threats}.

\ldawmsR's full data flow is illustrated in \myfig{fig-dfd-ldawmsr}. The proposal offers a few variations of the
self-escrow and authorization protocols. This figure uses the asymmetric key pair approach for Class key encryption in
the Secure Enclave, which creates strict hardware requirements. However, Savage offers a symmetric key variant that
achieves the same purposes. The asymmetric key variant was chosen for analysis because it is semantically simpler.

@@

When law enforcement needs access to a device, it first obtains a warrant to conduct the search. The warrant uniquely
identifies the suspect's mobile device. Law enforcement obtains the device, disassembles it to reveal the unexposed
\ac{EA} hardware interface, and begins the time-vaulted access request. After the lockout period is completed, the
device offers its Device ID only. The law enforcement agency's digital forensics department digitally signs the Device
ID and sends it to the device vendor, alongside the warrant and accompanying proof of legal authorization. The vendor's
access compliance department audits the legal documents and submits the digitally signed Device ID to a secure computing
environment such as a \ac{HSM}. The secure environment authenticates the digital signature, looks up the device's
private Device Seal Key and authorization token, and encrypts them with the law enforcement agency's public key. These
encrypted artifacts are returned to the agency, which decrypts and submits them to the device through the dedicated
interface. First, the authorization token is hashed and compared to the stored hash. If the hash matches, the device
accepts and uses the private Device Seal Key to decrypt the Class key, which was escrowed after being encrypted but the
public Device Seal Key. Once the Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for
forensic investigation.

This above process is a summary of the system in the \ldawmsr proposal. However, complete proposals must account for the
entire information lifecycle. In order to establish such a system, vendors and law enforcement must build mechanisms to
populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} depicts a \ac{DFD}
for these scenarios.

@@

The next step in threat modeling is to elicit threats based on the system's goals, threat actors, scope, and design.
Recall from \mysec{sec-crypto-basics} that security can be expressed in terms of authentication, integrity,
non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privileged
\cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
multiple categories, such as tampering for the effect of repudiation, are listed only once.

@@

Spoofing threats are listed in \mytab{table-spoofing}. Notably, \ii{S2} is Eran Tromer's attack \cite{tromer_2018} on
computer scientist Ray Ozzie's related \ac{DAR} \ac{EA} protocol called CLEAR \cite{ozzie_2018}. Savage acknowledges the
influence that CLEAR had on \ldawmsR, and specifically included a note that this proposal is technically vulnerable to
the same threat. This threat is not particularly severe, as the difficulty of such an attack is so high that it is
unlikely to be attempted. Even so, the mitigations listed here would also help.

@@

Tampering threats are listed in \mytab{table-tampering}. Note that execution of \ii{S2} above relies on a tampering
attack such as \ii{T1}.

@@

Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed, any attack
aimed at gaining unauthorized access to a device is an information disclosure threat. Most spoofing attacks aim to
disclose information intended for another party.

@@

The amount of risk \ldawmsr introduces depends on the number of suggested mitigations that are implemented. Some threats
introduce inherently little risk due to their difficult nature and low pay off. Others introduce greater risk but can be
reasonably mitigated. Some threats specifically target transparency, and therefore are difficult to address without
trust. Finally, there are some risks that cannot be safely mitigated. Each of these classes of threats are discussed in
turn.

\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low-risk. They
either they fail to produce useful results, can be reasonably mitigated, or are a threat that is already present in the
baseline depicted in \mysec{sec-basic-dar}. Although these threats are not fatal flaws to a design, it is important for
a design to consider and mitigate the risk they pose. Unless they appear in combination with other threats, they are do
not deserve more discussion.

\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
security, trustworthiness, and effectiveness of the proposal. However, they can be addressed through a set of related
requirements: require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device
to be physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities
the lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly entirely precludes certain attacks.
Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
Limiting the number of law enforcement facilities that can participate in the \ac{EA} system prevents LE impersonation
attacks. Together, these mitigations make the proposal more resistant to attack.

\ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
legal process and cooperation of several parties. Even in the U.S., the legal process can involve secret courts, secret
investigation, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is difficult to
achieve. A destructive disassembly requirement provides non-repudiation that the device was unlocked, if and only if one
has physical access to the device. Technically enforced transparency into device access requests and responses would
grant non-repudiation at a broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR}
\ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately,
building this feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider too much
transparency to be an anti-feature.

Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
data, the DID-DSK database. These threats create a large amount of risk that cannot be safely mitigated. To store so
much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
vendor needs to protect this database from the most advanced attackers in the world while regularly updating and
accessing it. \Acp{HSM} and robust security policies would make that a difficult task. The data will almost certainly
leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would use the
data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on their
enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
compliant device after dissembling it and waiting through the time vault.

@@

This chapter began with a list of goals by which \ac{DAR} \ac{EA} proposals can be measured, including security,
protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section \ldawmsr is
measured against each of these goals.

\paragraph*{Security} Security has already been discussed at length in this chapter. \ldawmsr introduces significant
risk in the DID-DSK database itself, but performs well in many other respects, especially if the recommended mitigations
are implemented. It goes from very hard to attack, when everything is working as intended, to simple to attack, when the
database is leaked. However, even when it becomes trivial to attack, the attackers are limited to devices in their
physical possession and cannot hide access from their victims.

\paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
authoritarian regimes. \ldawmsr does an excellent job of protecting against \ac{masssurv} due to its physicality. Any
power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior to a
leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After a
leak, these regimes could get into any device they want. If one lived in a country that was willing to take and destroy
their devices to read their contents without legal due process, this would be an unwelcome technology.

\paragraph*{Transparency} \ldawmsr provides transparency to the device user, but does not provide transparency to the
public. Due to destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed
using this method is irrefutable. However, this can only be confirmed if the device is available. Neither the user nor
the public can detect if a device is lawfully accessed under secretive conditions or is stolen and unlocked by
criminals. This leaves allows the potential for government abuse---though again with the caveat that physical possession
requirements mean it cannot be abused without a trace.

\paragraph*{Law enforcement utility} \ldawmsr performs a few tradeoffs between security and law enforcement utility. The
requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
authorized law enforcement laboratories each create barriers to usability. The physical possession requirement is far
too important for security and civil liberties to sacrifice for utility. Furthermore, law enforcement, and search and
seizure law generally, is geared towards physical evidence already. The time vault prevents against time-sensitive
threats, though destructive disassembly (not part of the original proposal) may be enough to cover these cases as well,
so the time vault could be decreased without significant impact. Limiting the number of laboratories able to use the
system reduces the risk of spoofing attacks; local law enforcement would have to cooperate with one of these special
laboratories, such as a regional \ac{FBI} office. (This really just shifts the spoofing threat up a level, though the
shift is appropriate.) In the end, law enforcement still has a reliable way to get into lawfully seized devices.
\ldawmsr upholds its end, but the government must fund the \ac{FBI} operation and cultivate cooperation between local
and federal agencies.
